---
title: "Report for Beth Israel Deaconess Medical Cetre"
date: "2022-10-26"
output: word_document
fontsize: 10pt
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(arsenal)
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(Epi)
require(grid) 

#set working directory
setwd("~/Documents/LSHTM/HDM/Project")


```

### Information on patient 42130 for the most recent hospital visit in which Simvastatin was prescribed

```{r, echo = FALSE,message=FALSE}

demographic <- read_csv("demographic.csv")
diagnoses <- read_csv("diagnoses.csv")
prescriptions <- read_csv("prescriptions.csv")

#convert the diagnoses and prescriptions to string to add to the table

diagnoses_list <- toString(as.list(diagnoses$diagnoses))
prescription_list <- toString(as.list(prescriptions$drug_name_generic))


#create df with prescriptions and diagnoses
diag_pres <- data.frame(diagnoses_list, prescription_list)


#rename the columns

names(demographic) <- c("Age", "Gender", "Ethnicity","Marital status", "Religion",
                        "Total time in hosptial","Time spent in ICU", "ICU unit(s) visited", "Ward ID")
names(diag_pres) <- c("Diagnoses", "Prescriptions")


#print out the table
knitr::kable(demographic, align = "clcccccc")
knitr::kable(diag_pres, align = "cc", "pipe") 


```

### Vital signs for patient 42130

```{r vitals, echo=FALSE, out.width="100%", fig.width = 10, fig.align='center',  message = FALSE}
vitals <- read_csv("time_series.csv")

#split input data frame into separate frames for ease of plotting
hr <- filter(vitals, itemid == 220045)
o2 <- filter(vitals, itemid == 220277)
resp_rate <- filter(vitals, itemid == 220210)
temp <- filter(vitals, itemid == 223761)
#use both aerterial and non-invasive bp as neither was taken over the entire icustay
bp1 <- filter(vitals, itemid == 220181)
bp2  <- filter(vitals, itemid == 220052)
bp <- rbind(bp1,bp2)

#plot the lines on a single plot
plot <-ggplot() +
          geom_line(data=hr, aes(x=charttime, y = valuenum, color = "Hear rate (bpm)" )) +
          geom_line(data=bp, aes(x=charttime, y = valuenum, color = "Blood pressure (mmHg)")) +
          geom_line(data=o2, aes(x=charttime, y = valuenum, color = "Blood oxygen (%)")) +
          geom_line(data=resp_rate, aes(x=charttime, y = valuenum, color = "Respiratory rate (per minute)")) +
          geom_line(data=temp, aes(x=charttime, y = valuenum, color = "Temperature (F)")) +
          labs(x="Chart time", y = "Vital value") +
          scale_color_manual(
                name = "Vital signs",
                breaks = c("Hear rate (bpm)", "blood pressure (mmHg)", "Blood oxygen (%)",
                                            "Respiratory rate (per minute)", "Temperature (F)"),
                values = c("Hear rate (bpm)" = "red", "blood pressure (mmHg)" = "blue",
                           "Blood oxygen (%)" = "green", "Respiratory rate (per minute)" = "orange",
                           "Temperature (F)" = "black"))
plot
    

```

### Summary of total time spent in ICU of 60-65 year olds with a cardiac device

Detailed below is information on patients of similar age to patient 42130, also with a cardiac device.


```{r cardiac, echo=FALSE, tab.align = 'left',  message = FALSE}

#read in the data
cardiac <- read_csv("icu_stay_time.csv")

#transform the d_in_icu field into yes or no

cardiac$d_in_icu <- factor(cardiac$d_in_icu, labels = c("No", "Yes"))

#create a stratified stat table using the Epi package

cardiac_summary <- stat.table(index=list("Gender"=gender, "Died in ICU?"=d_in_icu),
                             contents=list(Median=median(timeper_admit), LowerQuartile=quantile(timeper_admit, prob=0.25),
                                             UpperQuartile = quantile(timeper_admit, prob=.75),
                                             Min=min(timeper_admit), Max = max(timeper_admit)),
                              cardiac, margins=F)
names(cardiac_summary) <- c("Females who did NOT die in icu", 
                            "Males who did NOT die in icu",
                            "Females who DID die in icu", 
                            "Males who DID not die in icu")

knitr::kable(cardiac_summary)


```

Above time is given in fractional days. The data has been stratified by gender (M or F), and whether they have died in within 6 hours of being admitted or discharged from ICU (Yes or No).

### Summary of ICU stays for 60-65 year old patients with a cardiac device who died while in the ICU 

```{r, echo = FALSE, message=FALSE}
#set file path to wear the csv is stored
file_path <- "death_icu.csv"
death_icu <- read_csv(file_path)

#rename the column names
names(death_icu) <- c("ID", "M/F", "Age", "Stay time", "First ICD", "First 4 listed diagnoses")

#make the table
knitr::kable(death_icu)

```

### ICU stay times average

```{r icu_avg, echo= FALSE, out.width = "100%",, fig.algin = "center", message = FALSE, error=FALSE}

tot_avg <- read_csv("icu_average.csv")
card_avg <- read_csv("60-65_average.csv")
card_simv <- read_csv("60_65_cardiac.csv")

simv <- filter(card_simv, card_simv$simv==1)
no_simv <- filter(card_simv, card_simv$simv==0)

#remove the flag column 
simv = subset(simv, select = -simv )
no_simv = subset(no_simv, select = -simv )

#create a tibble with no stay time for the neo_nate patients
neo_nates <- tibble(careunit = c("NICU", "NWARD"), avg_staytime = c(0, 0) )

#bind the neo nates data to the data frames with 60-65 year old dat
card_avg <- rbind(card_avg, neo_nates)
simv <- rbind(simv, neo_nates)
no_simv <- rbind(no_simv, neo_nates)

#put care units in proper order
tot_avg <- tot_avg[order(tot_avg$careunit),]
card_avg <- card_avg[order(card_avg$careunit),]
simv <- simv[order(simv$careunit),]
no_simv <- no_simv[order(no_simv$careunit),]

names(tot_avg) <- c("careunit", "totstay")
names(card_avg) <- c("careunit", "cardavg")
names(simv) <- c("careunit", "simvstay")
names(no_simv) <- c("careunit", "nosimvstay")

# total <- merge(tot_avg, card_avg, by = "careunit")
total <- merge(total, simv, by ="careunit")
total <- merge(total, no_simv, by = "careunit")

ggplot(data= total, aes(x=careunit, y=c(totstay, cardavg, simvstay,nosimvstay)))
?gather
tot_plot <- ggplot(data=tot_avg, aes(x=careunit, y=avg_staytime, fill = careunit)) + 
            labs( x = "Care Unit", y = "Average stay length (days)", subtitle =  
                    "All patients") +
            geom_bar(stat="identity") +
            theme(axis.title.x=element_blank(),
                  axis.text.x=element_blank(),
                  axis.ticks.x=element_blank()) 
            
card_avg_plot <-ggplot(data=card_avg, aes(x=careunit, y=avg_staytime, fill = careunit)) + 
                labs( x = "Care Unit", y = "Average stay length (days)", subtitle =  
                        "60-65 with a cardiac device") + 
                geom_bar(stat="identity") +
                theme(axis.title.x=element_blank(),
                      axis.text.x=element_blank(),
                      axis.ticks.x=element_blank())

simv_plot <-  ggplot(data=simv, aes(x=careunit, y=avg_staytime, fill = careunit)) +
              labs( x = "Care Unit", y = "Average stay length (days)", subtitle =  
                      "60-65 with a cardiac device \n prescribed simvastatin")  +
              geom_bar(stat="identity") +
              theme(axis.title.x=element_blank(),
                    axis.text.x=element_blank(),
                    axis.ticks.x=element_blank())

no_simv_plot <- ggplot(data=no_simv, aes(x=careunit, y=avg_staytime, fill = careunit)) + 
                labs( x = "Care Unit", y = "Average stay length (days)", subtitle = 
                      "60-65 with a cardiac device \n not prescribed simvastatin") +
                geom_bar(stat="identity") + 
                theme(axis.title.x=element_blank(),
                      axis.text.x=element_blank(),
                      axis.ticks.x=element_blank())

#pu the plots together using ggarrange
figure <- ggarrange(tot_plot + rremove("ylab"), card_avg_plot + rremove("ylab"), simv_plot + rremove("ylab"), no_simv_plot + rremove("ylab"), 
          ncol = 2, nrow = 2, common.legend = TRUE)
annotate_figure(figure, left = textGrob("Time spent in ICU (days)", rot = 90, vjust = 1, gp = gpar(cex = 1.3)))

```
