#candidate number 221352

---
title: "Report for Beth Israel Deaconess Medical Cetre"
date: "2022-10-26"
output: word_document
fontsize: 10pt
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

## please uncomment any of the packages if you do not already have them installed
# install.packages("readr")
# install.packages("tidyverse")
# install.packages("ggplot2")
# install.packages("ggpubr")
# install.packages("grid")

library(readr)
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(Epi)
require(grid) 

#set working directory
setwd("~/Documents/LSHTM/HDM/Project")


```

### Information on patient 42130 for the most recent hospital visit in which Simvastatin was prescribed

```{r, echo = FALSE,message=FALSE}
path_to_file_demo <-"demographic.csv"
path_to_file_diag <- "diagnoses.csv"
demographic <- read_csv(path_to_file_demo)
diagnoses <- read_csv(path_to_file_diag)
prescriptions <- read_csv("prescriptions.csv")

#convert the diagnoses and prescriptions to string to add to the table

diagnoses_list <- toString(as.list(diagnoses$diagnoses))
prescription_list <- toString(as.list(prescriptions$drug_name_generic))


#create df with prescriptions and diagnoses
diag_pres <- data.frame(diagnoses_list, prescription_list)


#rename the columns

names(demographic) <- c("Age", "Gender", "Ethnicity","Marital status", "Religion",
                        "Total time in hosptial","Time spent in ICU", "ICU unit(s) visited", "Ward ID")
names(diag_pres) <- c("Diagnoses", "Prescriptions")


#print out the table
knitr::kable(demographic, align = "clcccccc")
knitr::kable(diag_pres, align = "cc", "pipe") 


```

### Vital signs for patient 42130

```{r vitals, echo=FALSE, out.width="100%", fig.width = 8, fig.align='center',  message = FALSE}
path_to_file_vital <- "time_series.csv"
vitals <- read_csv(path_to_file_vital)

#split input data frame into separate frames for ease of plotting
hr <- filter(vitals, itemid == 220045)
o2 <- filter(vitals, itemid == 220277)
resp_rate <- filter(vitals, itemid == 220210)
temp <- filter(vitals, itemid == 223761)
#use both aerterial and non-invasive bp as neither was taken over the entire icustay
bp1 <- filter(vitals, itemid == 220181)
bp2  <- filter(vitals, itemid == 220052)
bp <- rbind(bp1,bp2)

#plot the lines on a single plot
plot <-ggplot() +
          geom_line(data=hr, aes(x=charttime, y = valuenum, color = "Heart rate (bpm)" )) +
          geom_line(data=bp, aes(x=charttime, y = valuenum, color = "Blood pressure (mmHg)")) +
          geom_line(data=o2, aes(x=charttime, y = valuenum, color = "Blood oxygen (%)")) +
          geom_line(data=resp_rate, aes(x=charttime, y = valuenum, color = "Respiratory rate (per minute)")) +
          geom_line(data=temp, aes(x=charttime, y = valuenum, color = "Temperature (F)")) +
          labs(x="Chart time", y = "Vital value") +
          scale_color_manual(
                name = "Vital signs",
                breaks = c("Heart rate (bpm)", "Blood pressure (mmHg)", "Blood oxygen (%)",
                                            "Respiratory rate (per minute)", "Temperature (F)"),
                values = c("Heart rate (bpm)" = "red", "Blood pressure (mmHg)" = "blue",
                           "Blood oxygen (%)" = "green", "Respiratory rate (per minute)" = "orange",
                           "Temperature (F)" = "black"))
plot
    

```

### Summary of total time spent in ICU of 60-65 year olds with a cardiac device

Detailed below is information on patients of similar age to patient 42130, also with a cardiac device.


```{r cardiac, echo=FALSE,  message = FALSE}

#read in the data from correct csv file
path_to_file_staytime <- "icu_stay_time.csv"
cardiac <- read_csv(path_to_file_staytime)

#transform the d_in_icu field into yes or no

cardiac$d_in_icu <- factor(cardiac$d_in_icu, labels = c("No", "Yes"))

#create a stratified stat table using the Epi package

icu_summary <- stat.table(index=list("Gender"=gender, "Died in ICU?"=d_in_icu),
                             contents=list(Median=median(timeper_admit), LowerQuartile=quantile(timeper_admit, prob=0.25),
                                             UpperQuartile = quantile(timeper_admit, prob=.75),
                                             Min=min(timeper_admit), Max = max(timeper_admit)),
                              cardiac, margins=F)

cardiac_summary <- stat.table(index=list("Gender"=gender, "Died in ICU?"=d_in_icu),
                             contents=list(Median=median(hlos), LowerQuartile=quantile(hlos, prob=0.25),
                                             UpperQuartile = quantile(hlos, prob=.75),
                                             Min=min(hlos), Max = max(hlos)),
                              cardiac, margins=F)


knitr::kable(cardiac_summary, caption = "Summary of hosptial stay length (days)")
knitr::kable(icu_summary, caption = "Summary of total ICU stay length (days)")



```

The above data has been stratified by gender (M or F) and whether they have died in ICU, defined as dying within 6 hours of being admitted or discharged from ICU (Yes or No).

### Summary of ICU stays for 60-65 year old patients with a cardiac device who died while in the ICU 

```{r, echo = FALSE, message=FALSE}
#set file path to wear the csv is stored
file_path_death <- "death_icu.csv"
death_icu <- read_csv(file_path_death)

#rename the column names
names(death_icu) <- c("ID", "M/F", "Age", "Stay time", "First ICD", "Top 3 listed diagnoses")

#make the table
knitr::kable(death_icu)

```

### ICU stay times average

```{r icu_avg, echo= FALSE, out.width = "100%",, fig.algin = "center", message = FALSE, error=FALSE}

#set path to file to correct csv file
path_to_file_avg <- "icu_average.csv"
path_to_file_card <- "60-65_average.csv"
path_to_file_simv <- "simvastatin.csv"
path_to_file_no_simv <- "not_simvastatin.csv"

tot_avg <- read_csv(path_to_file_avg)
card_avg <- read_csv(path_to_file_card)
simv <- read_csv(path_to_file_simv)
no_simv <- read_csv(path_to_file_no_simv)

#create a tibble with no stay time for the neo_nate patients
neo_nates <- tibble(careunit = c("NICU", "NWARD"), avg_staytime = c(0, 0) )

#bind the neo nates data to the data frames with 60-65 year old dat
card_avg <- rbind(card_avg, neo_nates)
simv <- rbind(simv, neo_nates)
no_simv <- rbind(no_simv, neo_nates)

#put care units in proper order
tot_avg <- tot_avg[order(tot_avg$careunit),]
card_avg <- card_avg[order(card_avg$careunit),]
simv <- simv[order(simv$careunit),]
no_simv <- no_simv[order(no_simv$careunit),]

#rename the y-axis
names(tot_avg) <- c("careunit", "totstay")
names(card_avg) <- c("careunit", "cardavg")
names(simv) <- c("careunit", "simvstay")
names(no_simv) <- c("careunit", "nosimvstay")

#define the aggregate avg for each group, calculated in sql and transfered here

tot_agg <- 4.59
card_agg <- 3.59
simv_agg <- 3.99
no_simv_agg <- 3.50


#create four plots
tot_plot <- ggplot(data=tot_avg, aes(x=careunit, y=totstay, fill = careunit)) + 
            geom_hline(yintercept = tot_agg, linetype = "dashed") +
            labs( x = "Care Unit", y = "Average stay length (days)", subtitle =  
                    "All patients") +
            geom_bar(stat="identity") +
            theme(axis.title.x=element_blank(),
                  axis.text.x=element_blank(),
                  axis.ticks.x=element_blank()) 
            
card_avg_plot <-ggplot(data=card_avg, aes(x=careunit, y=cardavg, fill = careunit)) + 
                geom_hline(yintercept = card_agg, linetype = "dashed" ) +
                labs( x = "Care Unit", y = "Average stay length (days)", subtitle =  
                        "60-65 with a cardiac device") + 
                geom_bar(stat="identity") +
                theme(axis.title.x=element_blank(),
                      axis.text.x=element_blank(),
                      axis.ticks.x=element_blank())

simv_plot <-  ggplot(data=simv, aes(x=careunit, y=simvstay, fill = careunit)) +
              geom_hline(yintercept = simv_agg, linetype = "dashed") +
              labs( x = "Care Unit", y = "Average stay length (days)", subtitle =  
                      "60-65 with a cardiac device \n prescribed simvastatin")  +
              geom_bar(stat="identity") +
              theme(axis.title.x=element_blank(),
                    axis.text.x=element_blank(),
                    axis.ticks.x=element_blank())

no_simv_plot <- ggplot(data=no_simv, aes(x=careunit, y=nosimvstay, fill = careunit)) + 
                geom_hline(yintercept = no_simv_agg, linetype = "dashed"  ) +
                labs( x = "Care Unit", y = "Average stay length (days)", subtitle = 
                      "60-65 with a cardiac device \n not prescribed simvastatin") +
                geom_bar(stat="identity") + 
                theme(axis.title.x=element_blank(),
                      axis.text.x=element_blank(),
                      axis.ticks.x=element_blank())

#pu the plots together using ggarrange
figure <- ggarrange(tot_plot + rremove("ylab"), card_avg_plot + rremove("ylab"), simv_plot + rremove("ylab"), no_simv_plot + rremove("ylab"), 
          ncol = 2, nrow = 2, common.legend = TRUE) 
          
annotate_figure(figure, left = textGrob("Time spent in ICU (days)", rot = 90, vjust = 1, gp = gpar(cex = 1.3))) 
  

```

The dashed lines are the aggregated averages across the ICUs for each of the groups.